\subsection{Products and Coproducts}

\subsubsection*{Products}

\begin{definition}[Binary Product\index{Product}\label{def:product}]

	For a pair of objects $a, b$, a product candidate consists of an object $p$
	together with morphisms $p_1: P\to a$ and $p_2:P\to b$:
	% https://q.uiver.app/?q=WzAsMyxbMCwwLCJhIl0sWzIsMCwiYiJdLFsxLDAsInAiXSxbMiwwLCJwXzEiLDJdLFsyLDEsInBfMiJdXQ==
	\[\begin{tikzcd}[ampersand replacement=\&]
		a \& p \& b
		\arrow["{p_1}"', from=1-2, to=1-1]
		\arrow["{p_2}", from=1-2, to=1-3]
	\end{tikzcd}\]

	A product $(a\times b, \mathrm{fst}, \mathrm{snd})$ is a candidate such that
	for any candidate $(p, p_1, p_2)$ there exists a unique $h: p\to a\times b$
	that makes the following commute:
	\parencite{leinster:basic_category_theory}
	% https://q.uiver.app/?q=WzAsNCxbMCwxLCJhIl0sWzIsMSwiYiJdLFsxLDEsImFcXHRpbWVzIGIiXSxbMSwwLCJwIl0sWzIsMCwiXFxtYXRocm17ZnN0fSJdLFsyLDEsIlxcbWF0aHJte3NuZH0iLDJdLFszLDAsInBfMSIsMl0sWzMsMSwicF8yIl0sWzMsMiwiaCIsMix7InN0eWxlIjp7ImJvZHkiOnsibmFtZSI6ImRhc2hlZCJ9fX1dXQ==
	\[\begin{tikzcd}[ampersand replacement=\&]
		\& p \\
		a \& {a\times b} \& b
		\arrow["{\mathrm{fst}}", from=2-2, to=2-1]
		\arrow["{\mathrm{snd}}"', from=2-2, to=2-3]
		\arrow["{p_1}"', from=1-2, to=2-1]
		\arrow["{p_2}", from=1-2, to=2-3]
		\arrow["h"', dashed, from=1-2, to=2-2]
	\end{tikzcd}\]
\end{definition}

\begin{remark}
	Products are the dual of coproducts.
\end{remark}

\begin{definition}[Product as Limit\index{Product!as Limit}\label{def:product_limit}]

	More generally, for a discrete category $\I$, the functor $F:\I\to \C$ is an
	$\I$-indexed family of objects $(F(i))_{i\in \I_0}$ in $\C$. The product
	$\prod\limits_{i\in \I_0} F(i)$ is the limit of such a functor $F$.
	\parencite{leinster:basic_category_theory}
\end{definition}

\begin{remark}
	In the case where $\I=\underline{0}$ the product is denoted as $1$ and
	corresponds to a terminal object.
\end{remark}

\begin{remark}
	In the case where $\I=\underline{2}$, Definitions \ref{def:product} and
	\ref{def:product_limit} are equivalent.
\end{remark}

\begin{theorem}[Uniqueness of Product]
	The product of a collection of objects is unique up to isomorphism.

	\begin{proof}
		As the product of objects is a limit, and by Theorem \ref{thm:limit_iso}
		limits are unique up to isomorphism, then products are also unique up to
		isomorphism.
	\end{proof}
\end{theorem}

\subsubsection*{Coproduct}

\begin{definition}[Binary
Coproduct\index{Coproduct}\label{def:coproduct}]

	For a pair of objects $(a, b)$, a coproduct or sum candidate consists of an
	object $P$ together with morphisms $p_1: a\to p$ and
	$p_2:b\to p$:
	% https://q.uiver.app/?q=WzAsMyxbMCwwLCJhIl0sWzIsMCwiYiJdLFsxLDAsInAiXSxbMCwyLCJwXzEiXSxbMSwyLCJwXzIiLDJdXQ==
	\[\begin{tikzcd}[ampersand replacement=\&]
		a \& p \& b
		\arrow["{p_1}", from=1-1, to=1-2]
		\arrow["{p_2}"', from=1-3, to=1-2]
	\end{tikzcd}\]

	A coproduct $(a + b, \mathrm{left}, \mathrm{right})$ is a candidate such that
	for any candidate $(p, p_1, p_2)$ there exists a unique $h: a+b\to P$ that
	makes the following commute:
	\parencite{leinster:basic_category_theory}
	% https://q.uiver.app/?q=WzAsNCxbMCwwLCJhIl0sWzIsMCwiYiJdLFsxLDAsImErYiJdLFsxLDEsInAiXSxbMCwyLCJcXG1hdGhybXtsZWZ0fSJdLFsxLDIsIlxcbWF0aHJte3JpZ2h0fSIsMl0sWzAsMywicF8xIiwyXSxbMSwzLCJwXzIiXSxbMiwzLCJoIiwyLHsic3R5bGUiOnsiYm9keSI6eyJuYW1lIjoiZGFzaGVkIn19fV1d
	\[\begin{tikzcd}[ampersand replacement=\&]
		a \& {a+b} \& b \\
		\& p
		\arrow["{\mathrm{left}}", from=1-1, to=1-2]
		\arrow["{\mathrm{right}}"', from=1-3, to=1-2]
		\arrow["{p_1}"', from=1-1, to=2-2]
		\arrow["{p_2}", from=1-3, to=2-2]
		\arrow["h"', dashed, from=1-2, to=2-2]
	\end{tikzcd}\]
\end{definition}

\begin{remark}
	Coproducts are the dual of products.
\end{remark}

\begin{definition}[Coproduct as Colimit\index{Coproduct!as Colimit}\label{def:coproduct_limit}]
	For a discrete category $\I$, the functor $F:\I\to \C$ is an $\I$-indexed
	family of objects $(F(i))_{i\in \I_0}$ in $\C$. The coproduct or sum
	$\sum\limits_{i\in \I_0} F(i)$ is the colimit of such a functor $F$.
	\parencite{leinster:basic_category_theory}
\end{definition}

\begin{remark}
	In the case where $\I=\underline{0}$ the product is denoted as $0$ and
	corresponds to an initial object.
\end{remark}

\begin{remark}
	In the case where $\I=\underline{2}$, Definitions \ref{def:coproduct} and
	\ref{def:coproduct_limit} are equivalent.
\end{remark}

\begin{theorem}[Uniqueness of Coproduct]
	The coproduct of a collection of objects is unique up to isomorphism.

	\begin{proof}
		As the product of objects is a colimit, and by Theorem \ref{thm:colimit_iso}
		colimits are unique up to isomorphism, then coproducts are also unique up to
		isomorphism.
	\end{proof}
\end{theorem}